Validation of Visual Statistical Inference, Applied to Linear Models

### 2013-03-11 paper_jasa.tex
lineup numbers

The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated using a range of values of $\beta_2$ or slopes as shown in Table \ref{tbl:experiment_params}, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way. The actual data plot was randomly placed among these null data plots to produce the lineup.  \blue {Figure \ref{fig:test_category} is an example of one of these lineups. It was generated for $n$=300, $\beta_2$=10 and $\sigma$=12. The actual data plot location is the integer number nearest to $log_e(3269017)$.  For this lineup, 15 out of 16 observers picked the actual data plot. }

>>>>

The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated using a range of values of $\beta_2$ or slopes as shown in Table \ref{tbl:experiment_params}, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way. The actual data plot was randomly placed among these null data plots to produce the lineup.  \blue {Figure \ref{fig:test_category} is an example of one of these lineups. It was generated for $n$=300, $\beta_2$=10 and $\sigma$=12. The actual data plot location is the integer number nearest to $log_e(3269017)$. \hh{could you exchange that for a number that we can calculate without a calculator? $4^2-1$ would be completely fine - it's not supposed to be a real stumbling block, it's just supposed to not be immediately recognizable. }  For this lineup, 15 out of 16 observers picked the actual data plot. }
